---
title: "05_compute_completeness"
author: "Dexter H. Locke, PhD"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 0 set up: load libraries, custom functions, set defaults
```{r}

# load libraries
# packages we'll be using
packs <- c(
    'tidyverse'  # a must have
  , 'tidylog'    # prints out what was done in dplyr and tidr
  # , 'gbifdb' # GBIF
  , 'fst' # a faster table, makes outputs files much smaller, too.
  # , 'terra'
  , 'KnowBR'    # creates biodiversity estimates like completeness.
  , 'sf'        # spatial support
  , 'mapview'   # webmaps
  , 'janitor'   # cleans things up, also pipe-friendly cross-tabulations
  , 'tictoc'    # times things
  , 'beepr'     # makes noises
)


# check for all of the libraries, install if you don't have them
if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))  
}

# load them
vapply(packs, library, character.only = TRUE, logical(1), logical.return = TRUE, quietly = TRUE)



# set custom function for getting spatial data
see_sf <- function(){
# what's in memory that are sf - spatial features?
keep(eapply(.GlobalEnv, class),      # gets the objects in the global environment
     ~ any(str_detect(., "sf"))) %>% # selects elements with sf in them
    names(.) %>% as.character(.)     # my simple features
}

see_sf() -> sf_in_memory

# what are the spatial references of those SF classes?
mget(sf_in_memory) %>% purrr::map(~st_crs(.x)$epsg) %>% unlist() #%>% View()


# # get file size
# for(obj in ls()){message(obj); print(object.size(get(obj)), units='auto'); cat('\n')}; rm(obj)

# thanks Phil Donovan @philip_donovan
# https://www.spatialanalytics.co.nz/post/2018/04/01/fixing-st-par/
# Paralise any simple features analysis.
st_parallel <- function(sf_df, sf_func, n_cores, ...){

  # Create a vector to split the data set up by.
  split_vector <- rep(1:n_cores, each = nrow(sf_df) / n_cores, length.out = nrow(sf_df))

  # Perform GIS analysis
  split_results <- split(sf_df, split_vector) %>%
    parallel::mclapply(function(x) sf_func(x, ...), mc.cores = n_cores)


  # Define the output_class. If length is greater than two, then grab the second variable.
  output_class <- class(split_results[[1]])
  if (length(output_class) == 2){
    output_class <- output_class[2]
  }

  # Combine results back together. Method of combining depends on the output from the function.
  if (output_class == "matrix"){
    result <- do.call("rbind", split_results)
    names(result) <- NULL
  } else if (output_class == "sfc") {
    result <- do.call("c", split_results)
    result <- sf_func(result) # do.call combines the list but there are still n_cores of the geometry which had been split up. Running st_union or st_collect gathers them up into one, as is the expected output of these two functions.
  } else if (output_class %in% c('list', 'sgbp') ){
    result <- do.call("c", split_results)
    names(result) <- NULL
  } else if (output_class == "data.frame" ){
    result <- do.call("rbind", split_results)
  } else {
    stop("Unknown class. st_parallel only accepts the following outputs at present: sfc, list, sf, matrix, sgbp.")
  }

  # Return result
  return(result)
}


# Paralise any simple features analysis.
# https://www.spatialanalytics.co.nz/post/2017/09/11/a-parallel-function-for-spatial-analysis-in-r/
# define
st_par <- function(sf_df, sf_func, n_cores, ...){
  # Create a vector to split the data set up by.
  split_vector <- rep(1:n_cores, each = nrow(sf_df) / n_cores, length.out = nrow(sf_df))
  
  # Perform GIS analysis
  split_results <- split(sf_df, split_vector) %>%
    parallel::mclapply(function(x) sf_func(x, ...), mc.cores = n_cores)
  
  # Combine results back together. Method of combining depends on the output from the function.
  if (class(split_results[[1]]) == 'list' ){
    result <- do.call("c", split_results)
    names(result) <- NULL
    } else {
      result <- do.call("rbind", split_results)
      }
  # Return result
  return(result)
  }

# custom function for "Not In"
`%nin%` <- Negate(`%in%`)


# redlining colors
holc_pal <- c('#92BC6B' # green
              , '#92C7C9' # blue
              , '#E7DC6B' # yellow
              , '#E47D67' # red
              #, '#A9A9A9'
              ) # dark gray)

holc_pal_f<- c('#92BC6B' # green
              , '#92C7C9' # blue
              , '#E7DC6B' # yellow
              , '#E47D67' # red
              , '#A9A9A9'
              , '#000000')

# fixes mapview
mapviewOptions(fgb = FALSE)

sf::sf_use_s2(FALSE) # supresses error in some sf operations


data(adworld) # needed for KnowBPolygon
```


# 1 read in spatial data
## A holc polygons - is this even needed for right now?
```{r}

# sf::sf_use_s2(FALSE) # supresses error about invalid loops in 1212, 2851
# # https://stackoverflow.com/questions/68478179/how-to-resolve-spherical-geometry-failures-when-joining-spatial-data
# 
# tic(); (holc <- 
#           st_read("https://dsl.richmond.edu/panorama/redlining/static/fullDownload.geojson") |> 
#           # st_read('input_data/HOLC_shapefile/holc_ad_data.shp', as_tibble = TRUE) |> 
#           filter(!is.na(holc_grade) & holc_grade != 'E') %>%
#           st_cast('POLYGON') %>% # IMPORTANT
#           filter(!st_is_empty(.)) %>% 
#           st_make_valid(.) %>% 
#           rowid_to_column() %>% 
#           rowid_to_column(var = 'global_id') %>% 
#           # rowid_to_column(var = 'internal_id') %>% # TODO get this squeeky clean
#           mutate(  id = paste(state, city, holc_id, holc_grade, rowid, sep = '_')
#                  , city_state = paste0(city, ', ', state)
#                  , area_holc_km2 = as.double(st_area(.) / 1e+6)) %>% 
#   select(id, global_id, state, city, holc_id, holc_grade, city_state, area_holc_km2));toc() # < 5 seconds

# *** OR ***

# NOTICE !!! as_Spatial was failing because 'FL_Jacksonville_B3_B_1404' is a linstring, not a polygon
# F FL but go 'Gators!
(holc <- st_read('working_data/holc_polys_saved/holc_plys_2022-12-08 20-13-17.gpkg') |> 
  filter(st_is(geom, 'POLYGON')) |> 
  as_Spatial())



# map holc polygons and MSA
holc |> 
  # too many polygons at once.. filter by a state
  # filter(state == 'CA') |> 
  # filter(state == 'NY') |> 
  # filter(state == 'MA') |> 
  mapview(zcol =   'holc_grade'
                , col.regions = holc_pal)
  
```




TODO should we have dropped 'species' without genus and species before?

## B KnowBPolygon prep and runs

Make data ready for KnowBPolygon via grouping and selecting

### i fungi - test: work out the workflow on this subset of the data
```{r eval=FALSE, include=FALSE}

(
  fungi <- 
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_fungi_holc_2023-01-01.fst') |> 
    tibble() |> 
    select(genus, species, n_obs, X, Y, id, holc_grade) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id, holc_grade) |> 
    summarise(counts = sum(n_obs)) |> 
    select(genus_species, lon = X, lat = Y, counts, holc_grade)
)

# why the "Adding missing grouping variables: `id`" message?

# 2x checks
fungi |> tail()
fungi |> arrange(desc(counts))
fungi |> tabyl(id) |> tibble()
fungi |> tabyl(holc_grade)


# work only with data intersecting HOLC polygons per MSA (aka exclude the non-graded areas), much smaller
(
  fungi_holc <- # notice name
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_fungi_holc_2023-01-01.fst') |> 
    tibble() |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) # cosmetic reorder
)

# 2x checks
fungi_holc |> tail()
fungi_holc |> arrange(desc(counts))
fungi_holc |> tabyl(id) |> tibble()

(random_id <- fungi_holc |> sample_n(1) |> pull(id))
fungi_holc |> filter(id == random_id) # |> View()

# map
fungi_holc |> filter(id == random_id) |> st_as_sf(coords = c('lon', 'lat'), crs = st_crs(4326)) |> mapview() +
  mapview(holc[holc$id == random_id, ])
  # mapview(holc |> filter(id == random_id)) # old when sf

random_id <- c('CA_San Francisco_D3_D_1055', 'CA_San Francisco_D13_D_1048')

# dplyr::filter() + stringr::str_detect() by @mdsumner
detect_filter <- function(.data, string, pattern, negate = FALSE){
  tidylog::filter(.data, stringr::str_detect({{string}}, pattern, negate = negate))
  }

fungi_holc |> detect_filter(id, 'CA')

# trial test out
data(adworld)
tic(); KnowBPolygon(  #data = fungi_holc |> filter(id %in% random_id) |> data.frame() # testing on a few polys
                      data = fungi_holc |> detect_filter(id, 'CA') |> data.frame()    # testing on all of California
                     , format = 'A'
                     , estimator = 1
                     # , shape = holc |> filter(id %in% random_id) |> as_Spatial()     # sample polys
                     # , shape = holc |> detect_filter(id, 'CA') |> as_Spatial()   # testing on all of California
                     , shape = holc
                     , shapenames = 'id'
                     , save = 'CSV'
                     , dec = '.'
                     , jpg = FALSE
                     , Maps = FALSE
                     # , jpg = TRUE
                     # , Maps = TRUE
                     # , legend = TRUE
                     # , colscale = c( "#C8FFFFFF"
                     #                ,"#64FFFFFF"
                     #                ,"#00FFFFFF"
                     #                ,"#64FF64FF"
                     #                ,"#C8FF00FF"
                     #                ,"#FFFF00FF"
                     #                ,"#FFC800FF"
                     #                ,"#FF6400FF"
                     #                ,"#FF0000FF")
                     ); toc() # ~10 seconds for California, 8 without maps and figs, 12 seconds when fed ALL of holc polygons


```

### i fungi run - like above, but refined and for real
```{r eval=FALSE, include=FALSE}

# (
#   fungi <- 
#     read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_fungi_holc_2023-01-01.fst') |> 
#     tibble() |> 
#     select(genus, species, n_obs, X, Y, id, holc_grade) |> 
#     unite(genus_species, genus, species, sep = '_') |> 
#     filter(genus_species != 'NA_NA') |> # FIXME keep or not?
#     group_by(genus_species, X, Y, id, holc_grade) |> 
#     summarise(counts = sum(n_obs)) |> 
#     select(genus_species, lon = X, lat = Y, counts, holc_grade)
# )
# 
# # why the "Adding missing grouping variables: `id`" message? becuase of NAs in id?
# 
# # 2x checks
# fungi |> tail()
# fungi |> arrange(desc(counts))
# fungi |> tabyl(id) |> tibble()
# fungi |> tabyl(holc_grade)


# work only with data intersecting HOLC polygons per MSA (aka exclude the non-graded areas), *much* smaller
(
  fungi_holc <- # notice name
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_fungi_holc_2023-01-01.fst') |> 
    # tibble() |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
    data.frame()                                           # KnowBPolygon doesn't like tibble
)

# 2x checks
fungi_holc |> tail()
fungi_holc |> arrange(desc(counts))
fungi_holc |> tabyl(id) |> tibble()


setwd(paste0(getwd(), '/working_data/completeness/fungi_graded_only')) # FUNGUS
getwd()


tic(); KnowBPolygon(  data = fungi_holc 
                    , format = 'A'
                    , shape = holc
                    , shapenames = 'id'
                    , save = 'CSV'
                    , dec = '.'
                    , jpg = FALSE
                    , Maps = FALSE
                    # , jpg = TRUE
                    # , Maps = TRUE
                    # , legend = TRUE
                    # , colscale = c( "#C8FFFFFF"
                    #                ,"#64FFFFFF"
                    #                ,"#00FFFFFF"
                    #                ,"#64FF64FF"
                    #                ,"#C8FF00FF"
                    #                ,"#FFFF00FF"
                    #                ,"#FFC800FF"
                    #                ,"#FF6400FF"
                    #                ,"#FF0000FF")
                    ); toc() # 45 seconds for 13,804 points in 9839 polygons

# clean up
rm(fungi, fungi_holc)

# RESET THE WORKING DIRECTORY
setwd('../../../')
getwd()
```


### ii plantae
```{r eval=FALSE, include=FALSE}
# work only with data intersecting HOLC polygons per MSA (aka exclude the non-graded areas), much smaller


(
  plantae_holc <- # notice name
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_plantae_holc_2023-01-01.fst') |> 
    # tibble() |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
    data.frame()                                           # KnowBPolygon doesn't like tibble
)

# 2x checks
plantae_holc |> dim()
plantae_holc |> tail()
plantae_holc |> arrange(desc(counts))
plantae_holc |> tabyl(id) |> tibble()


setwd(paste0(getwd(), '/working_data/completeness/plantae_graded_only')) # FUNGUS
getwd()


tic(); KnowBPolygon(  data = plantae_holc 
                    , format = 'A'
                    , shape = holc
                    , shapenames = 'id'
                    , save = 'CSV'
                    , dec = '.'
                    , jpg = FALSE
                    , Maps = FALSE
                    ); toc() # 45 min for 918,255 points in 9839 polygons


# clean up
rm(plantae, plantae_holc)

# RESET THE WORKING DIRECTORY
setwd('../../../')
getwd()
```


### iii aves by year
```{r eval=FALSE, include=FALSE}
(aves_files <- 
    list.files('../biodiversity_in_a_concrete_jungle_data_too_big'
               , pattern = 'gbif_aves_holc' # MORE SPECIFIC THAN in script 03 and 04, bc it HAS TO BE
               , full.names = TRUE))

# # sloppy but works (all of the columns, growing a vector [ewwey])
# # good to know though for when we want to look at non-graded areas!
# tic(); for(i in 1:length(aves_files)){
#   print(i)
#   if(i == 1){ aves_holc <- read_fst(aves_files[i]) }
#   else      { aves_holc <- rbind(aves_holc, read_fst(aves_files[i])) }
#   }; toc(); beepr::beep() # ~2 mins
# 
# aves_holc |> dim() # 51,591,321 records!


# sloppy but works (all of the columns, growing a vector [ewwey])
tic(); for(i in 1:length(aves_files)){
  print(i)
  if(i == 1){
    
    aves_holc <- 
    read_fst(aves_files[i]) |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
    data.frame()                                           # KnowBPolygon doesn't like tibble (needed?)
    
    }
  else{
    
    aves_holc <- 
      rbind(aves_holc
            , read_fst(aves_files[i]) |> 
              filter(holc_grade != 'Not Graded') |> 
              select(genus, species, n_obs, X, Y, id) |> 
              unite(genus_species, genus, species, sep = '_') |> 
              filter(genus_species != 'NA_NA') |> # FIXME keep or not?
              group_by(genus_species, X, Y, id) |> 
              summarise(counts = sum(n_obs)) |> 
              ungroup() |> # FIXME why is needed at all?
              select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
              data.frame()                                           # KnowBPolygon doesn't like tibble
      )
    }
  }; toc(); beepr::beep() # ~36 SECONDS


# 2x checks
aves_holc |> dim() # 2,002,190 records
aves_holc |> tail()
aves_holc |> arrange(desc(counts))
aves_holc |> tabyl(id) |> tibble()


setwd(paste0(getwd(), '/working_data/completeness/aves_graded_only')) # FUNGUS
getwd()


tic(); KnowBPolygon(  data = aves_holc 
                    , format = 'A'
                    , shape = holc
                    , shapenames = 'id'
                    , save = 'CSV'
                    , dec = '.'
                    , jpg = FALSE
                    , Maps = FALSE
                    ); toc() # 1:33 for 2,002,190 points in 9839 polygons


# clean up
rm(aves, aves_holc)

# RESET THE WORKING DIRECTORY
setwd('../../../')
getwd()

```

### iv insecta
```{r eval=FALSE, include=FALSE}

# work only with data intersecting HOLC polygons per MSA (aka exclude the non-graded areas), much smaller


(
  insecta_holc <- # notice name
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_insecta_holc_2023-01-01.fst') |> 
    # tibble() |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
    data.frame()                                           # KnowBPolygon doesn't like tibble
)

# 2x checks
insecta_holc |> dim()
insecta_holc |> tail()
insecta_holc |> arrange(desc(counts))
insecta_holc |> tabyl(id) |> tibble()


setwd(paste0(getwd(), '/working_data/completeness/insecta_graded_only')) # FUNGUS
getwd()


tic(); KnowBPolygon(  data = insecta_holc 
                    , format = 'A'
                    , shape = holc
                    , shapenames = 'id'
                    , save = 'CSV'
                    , dec = '.'
                    , jpg = FALSE
                    , Maps = FALSE
                    ); toc() # 7 min for 121,489 points in 9839 polygons


# clean up
rm(insecta, insecta_holc)

# RESET THE WORKING DIRECTORY
setwd('../../../')
getwd()

```

### v mammalia
```{r eval=FALSE, include=FALSE}

# work only with data intersecting HOLC polygons per MSA (aka exclude the non-graded areas), much smaller

(
  mammalia_holc <- # notice name
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_mammalia_holc_2023-01-01.fst') |> 
    # tibble() |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
    data.frame()                                           # KnowBPolygon doesn't like tibble
)

# 2x checks
mammalia_holc |> dim()
mammalia_holc |> tail()
mammalia_holc |> arrange(desc(counts))
mammalia_holc |> tabyl(id) |> tibble()


setwd(paste0(getwd(), '/working_data/completeness/mammalia_graded_only')) # FUNGUS
getwd()


tic(); KnowBPolygon(  data = mammalia_holc 
                    , format = 'A'
                    , shape = holc
                    , shapenames = 'id'
                    , save = 'CSV'
                    , dec = '.'
                    , jpg = FALSE
                    , Maps = FALSE
                    ); toc() # ~1 min for 16,104 points in 9839 polygons


# clean up
rm(mammalia, mammalia_holc)

# RESET THE WORKING DIRECTORY
setwd('../../../')
getwd()

```

### vi reptilia (aka Squamata!)
```{r}


# work only with data intersecting HOLC polygons per MSA (aka exclude the non-graded areas), much smaller

(
  reptilia_holc <- # notice name
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_reptilia_holc_2023-01-02.fst') |> # dif date
    # tibble() |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
    data.frame()                                           # KnowBPolygon doesn't like tibble
)

# 2x checks
reptilia_holc |> dim()
reptilia_holc |> tail()
reptilia_holc |> arrange(desc(counts))
reptilia_holc |> tabyl(id) |> tibble()


setwd(paste0(getwd(), '/working_data/completeness/reptilia_graded_only')) # FUNGUS
getwd()


tic(); KnowBPolygon(  data = reptilia_holc 
                    , format = 'A'
                    , shape = holc
                    , shapenames = 'id'
                    , save = 'CSV'
                    , dec = '.'
                    , jpg = FALSE
                    , Maps = FALSE
                    ); toc() # ~24 seconds for 8,100 points in 9839 polygons


# clean up
rm(reptilia, reptilia_holc)

# RESET THE WORKING DIRECTORY
setwd('../../../')
getwd()

```

### vii amphibia
```{r}


# work only with data intersecting HOLC polygons per MSA (aka exclude the non-graded areas), much smaller

(
  amphibia_holc <- # notice name
    read_fst('../biodiversity_in_a_concrete_jungle_data_too_big/gbif_amphibia_holc_2023-01-01.fst') |> # dif date
    # tibble() |> 
    filter(holc_grade != 'Not Graded') |> 
    select(genus, species, n_obs, X, Y, id) |> 
    unite(genus_species, genus, species, sep = '_') |> 
    filter(genus_species != 'NA_NA') |> # FIXME keep or not?
    group_by(genus_species, X, Y, id) |> 
    summarise(counts = sum(n_obs)) |> 
    ungroup() |> # FIXME why is needed at all?
    select(genus_species, lon = X, lat = Y, counts, id) |> # cosmetic reorder
    data.frame()                                           # KnowBPolygon doesn't like tibble
)

# 2x checks
amphibia_holc |> dim()
amphibia_holc |> tail()
amphibia_holc |> arrange(desc(counts))
amphibia_holc |> tabyl(id) |> tibble()


setwd(paste0(getwd(), '/working_data/completeness/amphibia_graded_only')) # FUNGUS
getwd()


tic(); KnowBPolygon(  data = amphibia_holc 
                    , format = 'A'
                    , shape = holc
                    , shapenames = 'id'
                    , save = 'CSV'
                    , dec = '.'
                    , jpg = FALSE
                    , Maps = FALSE
                    ); toc() # ~16 seconds for 3,229 points in 9839 polygons


# clean up
rm(amphibia, amphibia_holc)

# RESET THE WORKING DIRECTORY
setwd('../../../')
getwd()
```


# 2 what is the relationship between size and speed for KnowBPolygon
```{r}

speed_size <- 
  tribble(
    ~size, ~speed,   ~taxon,
    13804,     42,    'fungi',
   918255,   2700,  'plantae', # 45 mins
  2002190,   7200,     'aves', # 1:30
   121489,    420,  'insecta', # 7 mins
    16104,     60, 'mammalia',
     8100,     24, 'reptilia',
     3229,     16, 'amphibia'  # (aka Squamata!)
  ) |> 
  mutate(duration = hms::as_hms(speed))

speed_size |> 
  ggplot(aes(size
             # , speed
             , duration
             )) + 
  geom_line() +
  geom_point(aes(color = taxon), size = 8) +
  # geom_step() + 
  theme_bw(16) +
  scale_x_continuous(labels = scales::comma) + 
  NULL


```

## End

# sand box

## find fucked up HOLC polygon(s) that wont convert sf to sp
```{r}

# sp::addAttrToGeom(x = holc$geometry)

holc |> as_Spatial() # I have no idea why this fails
holc |> filter(st_is(geometry, 'POLYGON')) |> as_Spatial() # I have no idea why this fails

# try to convert EACH polygon one at a time, to see where error arises.
for(i in holc$id){
  print(i)
  holc |> filter(id == i) |> as_Spatial()
}

# no issues. soo what's the deal?

# nest and map the as_Spatial function onto 
temp <- 
  holc |> 
  group_by(global_id) |> 
  nest() |> 
  mutate(sp = purrr::map(data, sf::as_Spatial, .progress = TRUE))

temp$sp[[1]]
temp$sp[[1]]@data
temp$sp[[1]] |> mapview()
rbind(temp$sp[[1]], temp$sp[[2]]) |> mapview()


for(i in temp$global_id){
  print(i)
  if(i==1){holc_sp <- temp$sp[[i]]} else {holc_sp <- rbind(holc_sp, temp$sp[[i]])}
  } # broke at feature 1403

# holc_sp
# all.true(holc_sp |> st_as_sf(), holc)

holc |> slice(1401 : 1407)
holc |> filter(id == 'FL_Jacksonville_B3_B_1404') |> mapview()
holc |> detect_filter(id, "FL") |> mapview()

```
